\Large \bf Cyclone: Replication Middleware for NVM Applications

### 2016-09-27 paper/paper.tex
updates.

\subsection*{Abstract}
The introduction of non-volatile directly attached memory in datacenters - in
the form of NVDIMMs and new memory technologies allows efficient data structures
to be built on non-volatile heaps that provide good performance coupled with
durability. However high availability remains a problem for these primarily
single machine solutions. While transactional interfaces built over RDMA capable
networking hardware offer a viable alternative, such hardware is often
unavailable in datacenters. Cyclone offers a solution to this problem by
providing transparent replication middleware that can be used with Intel's Non
Volatile Memory Library (NVML) to construct durable data-structures that map
keys to values.  Cyclone is designed around an optimized userspace networking
stack to reduce latency and can scale its consensus protocol across cores and
network ports to provide good throughput under load. We show that Cyclone can
replicate as many as 7 million updates a second using just 4 commodity 10 GigE
ports on a machine.

\section{Introduction}
Memory technology in datacenters is due to undergo a paradigm shift with the
increased use of directly addressable non-volatile memory both in the form of
battery backed non-volatile DIMMs~\cite{farm} as well as newer memory types
such as 3D XPoint~\cite{pmfs, bpfs}. In anticipation, libraries that allow
programmers to build durable data structures on non-volatile heaps have become
available~\cite{mnemosyne, nvheaps, cdds}  - an example of such a
library that we have used in this paper is Intel's non volatile memory library
(NVML)~\cite{nvml}. A missing piece however is a component to make such durable 
data structures highly available using replication across a commodity network
such as ethernet. Existing solutions that offer good performance are biased
towards pure transactions across RDMA with a dependence on external services
(such as zookeeper~\cite{zookeeper}) and application specific code for
fault recovery~\cite{farm, htm}.
This is a problem for programmers using NVML to construct durable and available 
data structures where one cannot apriori guarantee the availability of
RDMA (for e.g., when running on externally hosted environments or public
clouds).

Cyclone is prototype replication middleware that aims to demonstrate that one
can add fault tolerance transparently to a specific class of client server
applications written on top of libraries such as NVML. We specifically target
key-value maps in this paper and in that context make two key
contributions. First, we propose a simple API that enables server state to be
replicated without the programmer needing to worry about fault recovery. This
allows one to start by building a concurrent durable data structure using
libraries such as NVML and add availability at a later stage using Cyclone. The
second contribution is to show that one can achieve reasonably good replication 
performance using commodity ethernet via innovations in the networking stack and
by scaling replication with the application. This result should be a useful
design point for developers looking to deploy applications on non-volatile heaps
in commodity ethernet-based environments such as public clouds.

@@

\section{Replication}
Cyclone is designed to add fault tolerance to NVML client server applications
via replication -- a quorum of machines maintains \emph{equivalent} state.
Server state is queried and manipulated via RPC calls from clients.
Cyclone replicates the RPC call itself across replicas rather than replicating
every access (to the server state) made during execution of the RPC call.

Cyclone provides strongly consistent replication - every machine in the quorum
of replicas maintains a log of RPC calls. Cyclone treats the RPC call itself as
a variable sized opaque blob of data, leaving the application free to choose how
it wishes to marshall call related details and arguments. All machines agree on
the committed sequence of RPC calls in the log. We achieve this by running an
instance of the RAFT~\cite{raft} consensus protocol to keep the log of RPC calls
on different machines in synchronization. Specially flagged `read-only` RPC
calls are executed only on the leader replica and are not replicated.

\subsection{Decoupled Replication}
\label{sec:decouple}
The standard replicated state machine approach requires that commands (entries
in the log) be committed before they can be applied to the state machine --
i.e. in our case before the RPC call can be executed. In contrast, Cyclone
decouples execution of an RPC call on any replica from its replication.
The fact that RPC calls that modify state must themselves be executed
in a crash consistent transaction allows us to overlap execution with
replication. Normally, this would be a problem as replication might fail for a
number of reasons but primarily due to change of leadership and subsequent
rollback of previously appended (but not committed) log entries. To solve this
problem, Cyclone delays the actual commit of the failure atomic transaction
until after the replication is complete. In the event that replication fails,
the transaction is aborted and any changes to the non volatile heap are rolled
back. Figure~\ref{fig:async_rep} shows a pseudocode description of how the
runtime handles \emph{decoupled replication}.

@@

Decoupled replication provides good performance and we believe it would be
the common case for RPC calls replicated via Cyclone. However the program might 
on occasion require successful replication to be a precondition to exection.
A case where we find this necessary is
\textit{output commit}~\cite{output_commit} - where the RPC call has
external side effects (e.g., communication with other machines in the cluster)
and therefore crash consistency is insufficient to roll back the execution of
the RPC call in the event that replication fails. We therefore provide users the
ability to declare at the client side that an RPC call should execute
in a coupled manner at the server side. Cyclone would then ensure that the RPC
call is successfully replicated before it attempts execution on any of the
replicas.

@@

Cyclone separates the available CPU cores on a machine, in a configurable way,
into cores dedicated to work related to the (RAFT) consensus protocol and cores
decidated to running application code. Each RAFT core runs an \emph{independent}
instance of the RAFT protocol. RAFT cores pass work to the application cores
through lock-free double ended queues. A RAFT core dispatches an incoming RPC
call for execution simultaneously with sending it out for replication.

@@

In the event that the machine has multiple NICs, Cyclone spreads the network
demands and scales by binding each RAFT or application core to a specific NIC
and picking the necessary queues from that NIC.

\subsection{Zero Copy Batching}
Consensus protcols such as RAFT usually have a simple common case path. Once a
quorum leader is elected, the leader's primary job is to transmit new entries
appended to its persistent log to its followers. The primary job of a follower
is to receive these entries, append them to its persistent log and send
acknowledgments to the leader. An entry is considered committed when the leader
receives an acknowledgment for an entry from a majority of the quorum. Followers
receive intimation of committed entries as piggybacked information on subsequent
replication messages. The common case path at the leader lends itself to zero
copy since the same set of log entries is sent to all replicas.

In order to avoid an additional copy of the incoming RPC call contents from DRAM
to persistent memory, Cyclone allocates space for packet buffers on the non-volatile
heap. However, since packet data is cached, Cyclone ensures durability
(or persistence) of this data by explicitly flushing the dirty
cachelines~\cite{pmfs}. Cyclone maintains its log as a circular sequence of
pointers to packet buffers and therefore the append operation is completed by
adding a pointer to the received packet and flushing the cacheline containing
the pointer.

We exploit the fact that DPDK allows prepending data to a packet
without physically moving the data in a packet buffer. The received packet
contains only the RPC call from the client. We prepend a RAFT related header
(containing the previous log index, previous log term, leader's term and
leader's commit point~\cite{raft}) to the packet buffer itself. To this RAFT header 
we then prepend an IPv4 header. We assign pre-determined values to the
IP address fields in order to exploit 5-tuple steering at the receiving NIC for 
the placement of the incoming packet in the correct queue.
This composite packet buffer is then sent \emph{unchanged} to every replica.
To properly route the packet through switches, for each follower replica we
allocate a small buffer containing an ethernet header with appropriate addresses
and \emph{chain} the buffer in front of the prepared packet - thereby ensuring
that we do not need to change the contents of the packet itself.

Although the common case path in RAFT is straightforward, it still requires
hundreds of cycles (or a few microseconds) to process each log entry, which is
a significant amount in the face of sub 10 us network latency between nodes with
DPDK. We amortize this overhead when operating under load by
batching multiple RPC calls together and treating the batch as a single RAFT log
entry.  DPDK by itself can efficiently batch incoming packets under load,
therefore allowing the user to easily receive a burst of packets
(currently limited to 32) rather than a single packet for a polling call.
DPDK further allows chaining packet
buffers together and treating the chain as a single packet. This is convenient
for our zero copy approach and we chain packet mbufs together to form a single
packet. Since the entire chain is sent as a single packet when being replicated,
the length of a chain is limited by the size of the coalesced packet, that must
fit into an ethernet jumbo frame of approximately 9KB.

@@

An RPC call via Cyclone must traverse a specific RAFT core (and associated
consensus protocol instance) and specific application core. These details are
provided by the user when making the RPC call to the client side library
provided by Cyclone. The user therefore is aware of the number of execution
cores and RAFT instances and given complete control of \emph{RPC call steering}.
For most applications we envisage a simple modulo hashing scheme should suffice
for load balancing.

Cyclone provides at most once semantics. It supports a large but fixed (at
startup time) number of clients. Each RPC call from a client bears the client
number and a sequence number. \emph{Each execution core independently enforces
  that accepted RPC calls from the same client bear increasing sequence
  numbers}. This is done by storing the last executed sequence number for every
client in a per-core durable datastructure as part of the crash consistent
transaction wrapping the execution of the RPC call. Even in the presence of
fail-overs, this durable memory of sequence numbers seen from a client ensures
at most once semantics. In addition to guarding against repeated execution of
the same call, Cyclone also remembers the \emph{result} of the last executed RPC
call from every client. This is done to assist clients (particularly stateless
clients). Cyclone allows the client to query the last seen sequence number and
result for that RPC call from the quorum of servers.

@@

A durable hash table is simple to derive from a volatile hash table and readily
available in NVML as an example~\cite{nvml_hash}. To use that (non client-server)
example with Cyclone we added a simple RPC call format, manually marshalled and
demarshalled the few arguments to the RPC call in an RPC call handler we wrote -
a few tens of lines of code. Finally, we added a simple client driver program (a
few tens of additional lines of code).

@@

Our solution to this problem in Cyclone is to observe that replicas of the hash
table require \emph{per-key determinism} and not determinism across \emph{all}
operations. We therefore steer operations to the same key to the same RAFT core
and the same application core, by hashing the key to determine the RAFT core and
execution core. The actual configuration of the hash table therefore can be
different across different replicas - for example due to the insertion of two
keys mapping to the same hash bucket completing in different orders across
different replicas. However, since the results of insert, delete and lookup
operations for a key depend only on previous operations to the same key the
replicas provide identical results for committed operations.

@@

\subsection{Distributed Transactions}
A useful primitive for many distributed systems is the ability to run
distributed transactions - an atomic sequence of conditional updates spanning
multiple groups of replicas. Distributed transactions have two distinct facets:
atomicity and fault tolerance. Distributed transactions appear atomic with
respect to other atomic transactions and simple updates or reads by virtue of
standard concurrency control techniques. A common technique is strict two phase
locking: acquire all locks, update the locked objects and finally release all
the locks. We believe that NVM application programmers - particularly those
having experience with concurrent code should not find it hard to deal with
distributed concurrency. The hard part for most programmers is likely to be the
second aspect: fault tolerance. A distributed transaction spans many machines
and has a large footprint to clean up in the event of failure. Failure handling
is made more complex when considering cascading failures - further failures
while recovering from a failure. Cyclone was designed to allow primitives such
as distributed transctions to be added without the programmer needing to worry
about faults or write fault recovery code. As an exercise we show how to add
support for distributed transactions to our hash table.

@@

We guard against faults by executing the the entire transaction on a quorum of
machines via Cyclone. All machines in the quorum execute the transaction, making
calls to other quorums to execute steps in the two phase locking protocol. Since
execution of the RPC call leaves \emph{externally visible side effects} by
virtue of the RPC calls it makes we require output
commit~\cite{output_commit}. This is achieved if the execution is guaranteed to
complete once it starts. This is simply achieved by a flag in the RPC call that
turns off decoupled execution (Section~\ref{sec:decouple}). The flag causes
execution to start only after the RPC call has been successfully replicated.

@@

To ensure that the transaction is still guaranteed to complete, we use a
\emph{different} client identifier for each phase. Thus, until the transaction
completes and a new one is started the response for any RPC call continues to be
available. Figure~\ref{fig:dist_tx} shows in pseudocode the server side details
of a transaction. One initiated, the distributed transaction is guaranteed to
complete given quorum availabilty. However, if the client requires intimation of
completion, the final transaction state can be written to a distinguished
partition as part of the transaction itself.

We note that this design is not intended for performance, under the assumption
that distributed transactions will be infrequent compared to single key
operations.  On the other hand, it is worth comparing the lack of any need for
explicit fault recovery in Cyclone with fault recovery in systems systems that
explicity require such cleanup (~\cite{farm}, pp 59--63).  We believe that
such designs, which make availability easy for NVM applications are key to
increasing the adoption rate of NVM by datacenter operators.

@@

release:
 set client identifier to 3
 for each machine
   Update object
   If machine reponds RPC sequence too old
     Transaction completed by someone else

optional:
 write transaction result to <quorum_tx_result>
\end{verbatim}
}
\caption{Distributed Transactions}
\label{fig:dist_tx}
\end{figure}

\section{Evaluation}
We evaluate Cyclone against its design goals by using it for up to 3-way
replication in a cluster of twelve 48 core Xeon servers. Three servers run
Cyclone's server side code, while the remaining machines run multithreaded
clients. The server side machines are each
equipped with four commodity 82599-based 10 Gbit ethernet ports, while each
client machine has one 82599-based 10 Gbit ethernet port. All the machines are
connected through the same 10 Gbit switch. We use 8 cores on the
server machines for running 8 instance of RAFT, leaving 39 cores to run
application code. The last remaining core is reserved for management code. We
use NVML with the non volatile pool held in DRAM, which proxies for the
performance we would expect from NVDIMMs. We evaluate Cyclone in three stages.

First, we evaluate pure replication performance by replicating RPC
calls. Cyclone simply replicates the contents of the RPC call without executing
it and returns a response to the client when replication is complete.
Figure~\ref{fig:pure_rep} shows the results. We consider two cases, one where
the clients send an RPC call with no payload and one where they add a 512 byte
dummy payload. For the no payload case the bottleneck is the message handling
rate. We note that Cyclone can handle small message replication rates as high as
\emph{7 million messages a second}. In the case of RPC calls with 512 byte
payloads, the bottleneck is the available network bandwidth. To show this, we
also plot transmit throughput derived from the message by adding the rate at
which data is sent to replicas from all the 8 quorum leaders to the rate at
which responses are sent back to the clients.  It is interesting to note that
with 3 replicas Cyclone achieves \emph{more than 40 Gbps} of bandwidth. The
reason for this is that the quorum leaders can end up residing on different
machines. Although replication traffic is still limited by the receive bandwidth
at follower replicas, the responses to the clients, originating from different
machines have access to independent 4x10GigE ports, allowing total transmit
throughput to exceed 40 Gbps. Although not discussed in this paper, Cyclone's
control plane is designed to exploit this by trying to spread quorum leaders
across machines.

Also worth noting is that thanks to DPDK's low latency access to the NICs, three
way replication costs as little as 25 us of latency in the unloaded case. This
latency is measured at the client and therefore is four commodity 10 GigE
network hops through the switch, with each hop taking under 7 us including
network stack delays. These results should serve to illustrate that simple
commodity ethernet can indeed provide reasonable replication performance with a
carefully designed software stack that can exploit it.

@@

\section{Conclusion}
Cyclone allows developers to add availability to applications built on top of
libraries such as NVML and achieve good performance on commodity networking
hardware. A key question is whether Cyclone can be used for other applications
such as filesystems ? We believe that the requirement for scaling across
multiple replication quorums can be satisfied by applications where the scalable
commutativity rule~\cite{scalable_commutativity} applies to common case
primitives. Although our intitial release of Cyclone will be for
key value style applications, our future research is based on this hypothesis.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{acm}
\bibliography{paper}

>>>>

\subsection*{Abstract}
The introduction of non-volatile directly attached memory in datacenters - in
the form of NVDIMMs and new memory technologies allows efficient data structures
to be built on non-volatile heaps that provide good performance coupled with
durability. However high availability remains a problem for these primarily
single machine solutions. While transactional interfaces built over RDMA capable
networking hardware offer a viable alternative, such hardware is often
unavailable in datacenters. Cyclone offers a solution to this problem by
providing transparent replication middleware that can be used with Intel's Non
Volatile Memory Library (NVML) to construct durable data-structures that map
keys to values.  Cyclone is designed around an optimized userspace networking
stack to reduce latency. Further, it can scale its consensus protocol across
cores and network ports to provide good throughput under load. We show that
Cyclone can replicate as many as 7 million updates a second using just 8 cores
(for the consensus protocol) and 4 commodity 10 GigE ports on each replica.

\section{Introduction}
Memory technology in datacenters is due to undergo a paradigm shift with the
increased use of directly addressable non-volatile memory both in the form of
battery backed non-volatile DIMMs~\cite{farm} as well as newer memory types such
as 3D XPoint~\cite{pmfs, bpfs}. In anticipation, libraries that allow
programmers to build durable data structures on non-volatile heaps have become
available~\cite{mnemosyne, nvheaps, cdds} - an example of such a library that we
have used in this paper is Intel's non volatile memory library
(NVML)~\cite{nvml}. A missing piece however is a component to make such durable
data structures highly available using replication across a commodity network
such as ethernet. Existing solutions that offer good performance are dependent
on high performance RDMA networking hardware, with a dependence on external
services (such as zookeeper~\cite{zookeeper}) and application specific code for
fault recovery~\cite{farm, htm}.  This is a problem for programmers using NVML
where one cannot apriori guarantee the availability of such RDMA networking
hardware (for e.g., when running on externally hosted environments or public
clouds).

Cyclone is prototype replication middleware aiming to demonstrate that one
can add fault tolerance transparently to a specific class of client server
applications written on top of libraries such as NVML. We specifically target
key-value maps in this paper and in that context make two key
contributions. First, we propose a simple API that enables server state to be
replicated without the programmer needing to worry about fault recovery. This
allows one to start by building a concurrent durable data structure using
libraries such as NVML and add availability at a later stage using Cyclone. The
second contribution is to show that one can achieve reasonably good replication 
performance using commodity ethernet via innovations in the networking stack and
by scaling replication with the application. This result should be a useful
design point for developers looking to deploy applications on non-volatile heaps
in commodity ethernet-based environments such as public clouds.

@@

\section{Replication}
Cyclone is designed to add fault tolerance to NVML client server applications
via replication -- a quorum of machines maintains \emph{equivalent} state.
Server state is queried and manipulated via RPC calls from clients.  Cyclone
treats the RPC call itself as a variable sized opaque blob of data, leaving the
application free to choose how it wishes to marshall call related details and
arguments.

Cyclone replicates the RPC call across replicas rather than replicating
every access (to the server state) made during execution of the RPC call.
Cyclone provides strongly consistent replication - every machine in the quorum
of replicas maintains a log of RPC calls.  All machines agree on the committed
sequence of RPC calls in the log. We achieve this by running an instance of the
RAFT~\cite{raft} consensus protocol to keep the log of RPC calls on different
machines in synchronization. Specially flagged `read-only` RPC calls are
executed only on the leader replica and are not replicated.

\subsection{Decoupled Replication}
\label{sec:decouple}
The standard replicated state machine approach requires that commands (entries
in the log) be committed before they can be applied to the state machine --
i.e. in our case before the RPC call can be executed. This is because
replication of an accepted log entry can fail for a
number of reasons such as change of leadership and subsequent
rollback of previously appended (but not committed) log entries.

In contrast to other systems, Cyclone decouples execution of an RPC call on any
replica from its replication by executing the RPC call as soon as it is received
on any replica. We exploit the fact that RPC calls that modify state must
themselves be executed in a crash consistent transaction. Cyclone delays the
actual commit of the crash consistent transaction until after the replication is
complete. In the event that replication fails, the transaction is aborted and
any changes to the non volatile heap are rolled back. Figure~\ref{fig:async_rep}
shows a pseudocode description of how the runtime handles \emph{decoupled
  replication}.

@@

Decoupled replication provides good performance and we believe it would be the
common case for RPC calls replicated via Cyclone. However the program might on
occasion require successful replication to be a precondition to exection. A case
where we find this necessary is \textit{output commit}~\cite{output_commit} -
where the RPC call has external side effects (e.g., communication with other
machines in the cluster) and therefore crash consistency is insufficient to roll
back the execution of the RPC call in the event that replication fails. We
therefore provide users the ability to declare that an RPC call should execute
in a coupled manner. Cyclone would then ensure that the RPC call is successfully
replicated before it attempts execution on any of the replicas.

@@

Cyclone separates the available CPU cores on a machine, in a configurable way,
into cores dedicated to work related to the (RAFT) consensus protocol and cores
decidated to running application code. Each RAFT core runs an \emph{independent}
instance of the RAFT protocol - in conjunction with a similar instance on every
other replica. RAFT cores pass work to the application cores through lock-free
double ended queues. A RAFT core dispatches an incoming RPC call for execution
simultaneously with sending it out for replication.

@@

Cyclone can exploit multiple NICs on a machine by binding each RAFT or
application core to a specific NIC and picking the necessary queues from that
NIC.

\subsection{Zero Copy Batching}
Consensus protcols such as RAFT usually have a simple common case path. Once a
quorum leader is elected, the leader's primary job is to transmit new entries
appended to its persistent log to its followers. The primary job of a follower
is to receive these entries, append them to its persistent log and send
acknowledgments to the leader. An entry is considered committed when the leader
receives an acknowledgment for an entry from a majority of the quorum. Followers
receive intimation of committed entries as piggybacked information on subsequent
replication messages. The common case replication path at the leader lends
itself to \emph{zero copy} transmission to the replicas, since the same set of
log entries is sent to all replicas.

In the receive path, we avoid an additional copy of the incoming RPC call
contents from DRAM to persistent memory by allocating space for packet buffers
on the non-volatile heap. However, since packet data is cached, Cyclone must
ensure durability (or persistence) of this data by explicitly flushing the dirty
cachelines~\cite{pmfs}. Cyclone maintains its log as a circular sequence of
pointers to packet buffers and therefore the append operation is completed by
adding a pointer to the received packet and flushing the cacheline containing
the pointer.

We exploit the fact that DPDK allows prepending data to a packet
without physically moving the data in a packet buffer. The received packet
contains only the RPC call from the client. We prepend a RAFT related header
(containing the previous log index, previous log term, leader's term and
leader's commit point~\cite{raft}) to the packet buffer itself. To this RAFT header 
we then prepend an IPv4 header. We assign pre-determined values to the
IP address fields in order to exploit 5-tuple steering at the receiving NIC for 
the placement of the incoming packet in the correct queue.
This composite packet buffer is then sent \emph{unchanged} to every replica.
To properly route the packet through switches, for each follower replica we
allocate a small buffer containing an ethernet header with appropriate addresses
and \emph{chain} the buffer in front of the prepared packet - thereby ensuring
that we do not need to copy the contents of the packet itself.

Although the common case path in RAFT is straightforward, it still requires
hundreds of cycles (or a few microseconds) to process each log entry, which is a
significant amount in the face of sub 10 us network latency between nodes with
DPDK. We amortize this overhead when operating under load by batching multiple
RPC calls together and treating the batch as a single RAFT log entry.  DPDK by
itself can efficiently batch incoming packets under load, therefore allowing the
user to easily receive a burst of packets (currently limited to 32) rather than
a single packet for a polling call.  DPDK further allows chaining packet buffers
together and treating the chain as a single packet. We chain these packet
buffers together to form a single packet. Since the entire chain is sent as a
single packet when being replicated, the length of a chain is limited by the
size of the coalesced packet, that must fit into an ethernet jumbo frame of
approximately 9KB.

@@

An RPC call via Cyclone must traverse a specific RAFT core (and associated
consensus protocol instance) and specific application core. These details are
provided by the user when making the RPC call to the Cyclone client side
library. The user therefore is aware of the number of execution cores and RAFT
instances and given complete control of \emph{RPC call steering}. For most
applications we envisage a simple modulo hashing scheme should suffice for load
balancing, something that we plan to later automate in Cyclone.

Cyclone provides at most once semantics. It supports a large but fixed (at
startup time) number of clients. Each RPC call from a client bears the client
number and a sequence number. \emph{Each execution core independently enforces
  that accepted RPC calls from the same client bear increasing sequence
  numbers}. This is done by storing the last executed sequence number for every
client in a per-core durable data structure as part of the crash consistent
transaction wrapping the execution of the RPC call. Even in the presence of
fail-overs, this durable memory of sequence numbers seen from a client ensures
at most once semantics. In addition to guarding against repeated execution of
the same call, Cyclone also remembers the \emph{result} of the last executed RPC
call from every client. This is done to assist clients - particularly stateless
clients. Cyclone allows the client to query the last seen sequence number and
result for that RPC call from any application core on the quorum of servers.

@@

A durable hash table is simple to derive from a volatile hash table and readily
available in NVML as an example~\cite{nvml_hash}. To use that (non
client-server) example with Cyclone we added a simple RPC call format, manually
marshalled and demarshalled the few arguments to the RPC call in an RPC call
handler - a few tens of lines of code. Finally, we added a simple client driver
program (a few tens of additional lines of code).

@@

Our solution to this problem in Cyclone is to observe that replicas of the hash
table require \emph{per-key determinism} and not determinism across \emph{all}
operations. We therefore steer operations to a particular key to the same RAFT
core and the same application core (using a hash on the key).  The actual
configuration of the hash table therefore can be different across different
replicas - for example due to the insertion of two keys mapping to the same hash
bucket completing in different orders across different replicas. However, since
the results of insert, delete and lookup operations for a key depend only on
previous operations to the same key the replicas provide identical results for
committed operations. \emph{In other words, we exploit the fact that operations to
different keys in a map are commutative.}

@@

\subsection{Distributed Transactions}
A useful primitive for many distributed systems is the ability to run
distributed transactions - an atomic sequence of conditional updates spanning
multiple groups of replicas. Distributed transactions have two distinct facets:
atomicity and fault tolerance. Distributed transactions appear atomic with
respect to other atomic transactions and simple updates or reads by virtue of
standard concurrency control techniques. A common technique is strict two phase
locking: acquire all locks, update the locked objects and finally release all
the locks. We believe that NVM application programmers - particularly those
with experience in concurrency should not find it hard to deal with
distributed concurrency. The hard part for most programmers is likely to be the
second aspect: fault tolerance. A distributed transaction spans many machines
and has a large footprint to clean up in the event of failure. Failure handling
is made more complex when considering cascading failures - further failures
while recovering from a failure. Cyclone was designed to allow primitives such
as distributed transctions to be added without the programmer needing to worry
about faults or writing fault recovery code. As an exercise we show how to add
support for distributed transactions to our hash table.

@@

We guard against faults by executing the the entire transaction on a quorum of
machines via Cyclone. All machines in the quorum execute the transaction, making
calls to other quorums to execute steps in the two phase locking protocol. Since
execution of the RPC call leaves \emph{externally visible side effects} by
virtue of the further RPC calls it makes, we require output
commit~\cite{output_commit}. This is achieved if the execution is guaranteed to
complete once it starts. Hence, for distributed transactions, we turn off
decoupled execution (Section~\ref{sec:decouple}).

@@

To ensure that the transaction is still guaranteed to complete, we use a
\emph{different} client identifier for each \emph{phase}. Thus, until the
transaction completes and a new one is started the response for any RPC call
continues to be available. Figure~\ref{fig:dist_tx} shows in pseudocode the
server side details of a transaction. Once initiated, the distributed transaction
is guaranteed to complete given availability of a quorum.
%However, if the client
%requires intimation of completion, the final transaction state can be written to
%a distinguished partition as part of the transaction itself.

Distributed transactions built around at most once semantics are not intended
for performance. On the other hand, at most once semantics allows users of
Cyclone to avoid writing explicit fault recovery code as compared to fault
recovery in systems that explicity require such cleanup (~\cite{farm}, pp
59--63). Under the assumption that distributed transactions will be infrequent
compared to single key operations, such simplicity is key to increasing the
adoption of NVM among datacenter operators concerned about availability.

@@

release:
 set client identifier to 3
 for each machine
   Update object
   If machine reponds RPC sequence too old
     Transaction completed by someone else
\end{verbatim}
}
\caption{Distributed Transactions}
\label{fig:dist_tx}
\end{figure}

\section{Evaluation}
We evaluate Cyclone against its design goals by using it for up to 3-way
replication in a cluster of twelve 48 core Xeon servers. Three servers run
Cyclone's server side code, while the remaining machines run multithreaded
clients. The server side machines are each equipped with four commodity
82599-based 10 Gbit ethernet ports, while each client machine has one
82599-based 10 Gbit ethernet port. All the machines are connected to the same 10
Gbit switch. We use 8 cores on the server machines for running 8 instances of
RAFT and 39 cores to run application code. The last remaining core is reserved
for management code. We use NVML with the non volatile pool held in DRAM, which
proxies for the performance we would expect from NVDIMMs. We evaluate Cyclone in
three stages.

First, we evaluate pure replication by having Cyclone simply replicate the
contents of the RPC call without executing it and return a response to the
client when replication is complete.  Figure~\ref{fig:pure_rep} shows the
results. We consider two cases, one where the clients send an RPC call with no
payload and one where they add a 512 byte dummy payload. For the no payload case
the bottleneck is the message handling rate. We note that Cyclone can handle
small message replication rates as high as \emph{7 million messages a
  second}. In the case of RPC calls with 512 byte payloads, the bottleneck is
the available network bandwidth. To show this, we also plot aggregate transmit
throughput derived from these results by adding the rate at which data is sent
to replicas from all the 8 quorum leaders to the rate at which responses are
sent back to the clients.  With 3 replicas Cyclone shows \emph{slightly more
  than 40 Gbps} of aggregate transmit bandwidth. The reason for this is that the
quorum leaders can end up residing on different machines. Although replication
traffic is still limited by the receive bandwidth at follower replicas, the
responses to the clients, originating from different machines have access to
independent 4x10GigE ports, allowing total transmit throughput to exceed 40
Gbps. Although not discussed in this paper, Cyclone's control plane is designed
to exploit this by trying to spread quorum leaders across machines.

Also worth noting is that thanks to DPDK's low latency access to the NICs, three
way replication costs as little as 25 us of latency in the unloaded case. This
latency is measured at the client and is therefore four commodity 10 GigE
network hops through the switch, with each hop taking under 7 us including
network stack delays. These results should serve to illustrate that simple
commodity ethernet can indeed provide reasonable replication performance with a
carefully designed software stack that can exploit it.

@@

\section{Conclusion}
Cyclone allows developers to add availability to applications built on top of
libraries such as NVML and achieve good performance on commodity networking
hardware. A key question is whether Cyclone can be used for other applications
such as filesystems ? \emph{We believe that the requirement for scaling across
multiple replication quorums can be satisfied by applications where the scalable
commutativity rule~\cite{scalable_commutativity} applies to common case
primitives.} Although our intitial release of Cyclone will be for
key value style applications, our future research is based on this hypothesis.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{acm}
\bibliography{paper}
